---
title: "Activity 2 Presentation: Exploring Bivariate Data Relationships"
author: "Your Name"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    latex_engine: xelatex
  html_document: 
bibliography: references.bib
---

## Introduction

This presentation explores the relationship between two variables from a real-world dataset. Our objective is to determine if there exists a causal, no, or spurious relationship between these variables. This analysis is inspired by the Harvard Business Review article [@HBR2015] about the Spurious Correlation website [@VigenSC].

## Data Preparation

The dataset used in this analysis was sourced from [describe data source here]. The variables of interest are [Variable 1] and [Variable 2], chosen because [briefly explain why].

```{r load-data}
# Your R code to load (and preview) the dataset
```

## Data Visualization

```{r base-plot}
# Create a scatterplot of your variables + any other plots you find necessary
```

## Discussion

The plot(s) generated reveals [describe the overall pattern]. This pattern suggests that the relationship between Variable 1 and Variable 2 is [causal/no/spurious].

### Relationship Analysis

-   **Form of Relationship**: The relationship appears to be [linear/non-linear/etc.], indicating [explain].

-   **Causality**: Given the nature of the data, it is [likely/unlikely] that this relationship is causal because [reason].

-   **Alternative Explanations**: It is also possible that [other variables/confounding factors] could explain this relationship.

### Conclusion

In conclusion, the analysis of [Variable 1] and [Variable 2] provides [summarise findings]. While definitive conclusions about causality cannot be drawn without further information, the data suggests [final thoughts].

## References (optional)

-   \<To display the references, you need to make sure that `references.bib` is in the same directory as this R Markdown file. The `bibliography: references.bib` line in the header tells R Markdown where to find your bibliography file.\>

-   \<To add references, simply add entries of the same format in `references.bib`.\>

<!-- -->

-   \<To remove the references, you simply need to delete the line `bibliography: references.bib` in the header, and remove [@HBR2015] and [@VigenSC] in the introduction section.\>
