---
title: "ST117 Lab 9 Workbook"
author: ""
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    latex_engine: xelatex
  html_document: 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Linear Regression Model

Consider the regression model: 
$$
y_i= \alpha + \beta x_i + \epsilon_i
$$

where $\epsilon_i \sim \text{N}(0,\sigma^2)$ for $i = 1,\ldots,n$. Here, y represents heights(in cm) and x represents weights(in kg).

Given a sample of heights and weights,

```{r}
x<-c(82.62954, 66.73767, 83.29799, 82.72429, 74.14641, 54.60050, 60.71433,
     67.05280, 69.94233, 94.04653)
y<-c(202.5511,166.1328,194.1981,197.4564,181.9675,146.2233,160.5469, 166.2354,
     178.0746, 213.0961)
```

(a) Find the maximum likelihood estimates for $\alpha$ and $\beta$ based on the given sample.

```{r}
#TODO: write your codes here
```

(b) Plot a scatter plot of the sample and include the estimated regression line using the MLE of $\alpha$ and $\beta$. Providing the exact regression equation $y=50+1.8*x+\epsilon$, where $\epsilon \sim N(0,5^2)$, add this line into the plot.

```{r}
#TODO: write your codes here
```

(c) Generate 100 simulations and calculate 100 values of $\hat{\alpha}$ and $\hat{\beta}$ using the regression $y=50+1.8*x+\epsilon$, where $\epsilon \sim N(0,5^2)$ and $x \sim N(70,10^2)$. Calculate the variance of the MLE $(\hat{\alpha},\hat{\beta})$ using R and the formulas provided in Tuesday-1 Week 8 Lecture notes:

$$
\begin{aligned}
& \operatorname{Var}(\hat{\beta})=\frac{\sigma^2}{\sum_{i=1}^n\left(x_i-\bar{x}\right)^2} \\
& \operatorname{Var}(\hat{\alpha})=\sigma^2\left(\frac{1}{n}+\frac{\bar{x}^2}{\sum_{i=1}^n\left(x_i-\bar{x}\right)^2}\right)
\end{aligned}
$$

```{r}
#TODO: write your codes here
```

(d) Find the estimators' variances as the sample size $n$ increases, considering $n = 100, 1000, 10000$.

```{r}
#TODO: write your codes here
```

## 2. Diagnostic Check

(a) Consider `women` dataset which contains average heights and weights for American Women. Create a scatter plot of this data to examine the relationship between heights and weights. Is there evidence of a linear relationship?

```{r echo=TRUE}
 #access the dataset
data(women) #access the dataset

head(women)
```

```{r}
#TODO: write your codes here
```

(b) Fit a linear model to the data and perform a regression diagnostics check, including standard linear model diagnostics.

```{r}
#TODO: write your codes here
```

