---
title: "ST117 Lab 6 Workbook"
author: ""
date: "2024-02-16"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Bernoulli Distribution

Suppose we observe data $\vec{x}=(1,0,0,1,0,1,0,0,1,1)$ where each element $X_i$ follows a Bernoulli distribution with an unknown success probability $p$.

(a) Find the likelihood function $L(p;\vec{x})$ which represents the joint pdf function of $p$ given the observed data $\vec{x}$ and define it as a function in R. Calculate the value of this function at $p=0.1$.

```{r}
#TODO: write your codes here
```



(b) Plot the likelihood function for $p\in[0,1]$ and use the `optimize()` function to find the point that maximizes $L(p;\vec{x})$. Add a vertical line to the plot to indicate this maximum point.

```{r}
#TODO: write your codes here
```

(c) Calculate the log-likelihood function $l(p;\vec{x})$ and define this function in R. Plot the log-likelihood function. Verify that $p=0.5$ maximizes the log-likelihood function by setting its first derivative equal to zero and ensuring that the second derivative is negative.

```{r}
#TODO: write your codes here
```

## 2. Binomial Distribution
Given that $X \sim \text{Bin}(n,p)$ and observed that $n=10$ and $x=3$.

(a) Define the likelihood function in R and calculate it at $p=0.1$.
```{r}
#TODO: write your codes here
```

(b) Plot the likelihood function for $p\in[0,1]$ and calculate the maximum likelihood estimate. Add a vertical line to the plot to indicate this maximum point.
```{r}
#TODO: write your codes here
```

```{r}
#TODO: write your codes here
```


## 3. Normal Distribution
Given the heights(in cm) of a random sample of 20 students:

182, 154, 147, 150, 164, 177, 169, 173, 160, 173, 170, 160, 178, 175, 154, 179, 168, 188, 172, 162

We assume that the heights of students follow a normal distribution with unknown mean $\mu$ and standard deviation $\sigma$. 

(a) Determine the maximum likelihood estimates of $\mu$ and $\sigma$ based on the given sample.

```{r}
#TODO: write your codes here
```


(b) Calculate the log-likelihood of these height estimates.

```{r}
#TODO: write your codes here
```


