---
title: "ST117 Lab 7 Workbook"
author: ""
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: 
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tinytex)
```

## 1. Linear Regression Model

Generate a sample from the linear regression model:

$$
Y = 2X + \varepsilon, \ \varepsilon \sim \mathcal{N}(0,5)
$$

```{r}
#TODO: write your codes here
```

(a) Find the regression line to predict $\hat{y}$ by $x$, i.e. find $\hat{a}_1$, $\hat{b}_1$ such that 

$$
Y = \hat{b}_1 X + \hat{a} _1
$$

manually using the covraiance and correlation estimators. Then verify the values using the `lm` function in R.

```{r}
#TODO: write your codes here
```

(b) Find the regression line to predict $\hat{x}$ by $y$, i.e. find $\hat{a}_2$, $\hat{b}_2$ such that 

$$
X = \hat{b}_2 Y + \hat{a}_2 
$$

manually using the covraiance and correlation estimators. Then verify the values using the `lm` function in R.

```{r}
#TODO: write your codes here
```

(c) Plot the data along with the two regression lines. Which regression line appears to be better?

```{r}
#TODO: write your codes here
```

(d) The Least Squared Estimation can also help us derive the estimates of $\hat{a}$ and $\hat{b}$. We just need to find the value of $a$ and $b$ that minimizes the following expression:
$$
\sum_{i=1}^{n} \Big( y_i-(a+bx_i) \Big)^2
$$
Hint: Use the `optim` function.

```{r}
#TODO: write your codes here
```

## 2. Mean Squared Error
Consider a random sample $X_1,X_2,\ldots,X_n$ from a normal distribution with mean $\mu$ and variance $\sigma^2$. We have the following two estimators of the population variance obtained from the given sample information. From our previous lab, we derived the Maximum Likelihood Estimator(MLE) of the population variance as:
$$
1.\ \hat{\sigma}^2=\frac{1}{n}\sum_{i=1}^{n}(X_i-\overline{X})^2.
$$
Additionally, we have the sample variance given by:
$$
2.\ s^2=\frac{1}{n-1}\sum_{i=1}^{n}(X_i-\overline{X})^2,
$$
where $\overline{X}$ represents the sample mean.

(a) Simulate a sample of 10 data from $\text{N}(160,10^2)$. Calculate the values of these two estimators and their Mean Squared Error(MSE).

```{r}
#TODO: write your codes here
```

(b) Try  sample sizes of 100, 1000 and 10000, and show the results in a table. Can you provide an interpretation?

```{r}
#TODO: write your codes here
```