---
title: "ST117 Lab 3 Workbook"
author: ''
date: '2024-01-28'
output:
  html_document:
    theme: cosmo
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Simulating and Visualizing Binomial Distribution

(a) Generate a random sample of 1000 observations from a binomial distribution with parameters $n=20$ and $p=0.5$, then create a histogram to visualize the distribution.

```{r}
# TODO: generate binomial samples
```

```{r}
# TODO: write down your codes to plot the histogram
```

(b) Calculate the mean and standard deviation of the generated data. Compare these values with the theoretical mean and variance.

```{r}
# TODO: calculate mean and variance
```

(c) Flipping a coin 20 times, where it has 50% chance of getting a head. The total number of head $X$ follows a Binomial distribution with $n=20$ and $p=0.5$. Find the probability of getting between 5 and 12 heads when tossing 20 coins, which is sum of the orange bars.

```{r echo=FALSE, warning=FALSE}
# this chunk of code generates the plot demonstration, can ignore
# use ggplot to show the pdf of a binomial distribution  with shaded areas.
library("ggplot2")
n<-20
p<-0.5
# Generate values for the binomial distribution
x <- seq(0, n, by = 1)
binomial_pmf <- dbinom(x, size = n, prob = p)


# Create a data frame for ggplot

df <- data.frame(x = x, binomial_pmf = binomial_pmf)
df$block_color <- "grey70"
df[6:13,]$block_color <- "orange"

# Plot the binomial distribution
ggplot(df, aes(x = x, y = binomial_pmf,fill=block_color)) +
  geom_histogram(stat = "identity",binwidth=1,col = "black") +
  labs(title = "Binomial(20,0.5)",
       x = "Number of Successes",
       y = "Probability") +
  theme_minimal()+
  
  scale_fill_manual(values=c("grey70", "orange"))

```

```{r}
# TODO: calculate the probability
```

## 2. Calculating Probabilities from the Normal Distribution

(a) Suppose $X \sim \text{N}(0,1)$, find the value $m$ (green point) such that $P(X\leq m)=0.95$.

```{r echo=FALSE, warning=FALSE}
# this chunk of code generates the plot demonstration, can ignore
# use ggplot2 to plot a pdf of a normal distribution
library(ggplot2)

# Set parameters for the normal distribution
mean_value <- 0
sd_value <- 1

# Generate data for the normal distribution
x <- seq(-3, 3, 0.01)
pdf_normal <- dnorm(x, mean = mean_value, sd = sd_value)

# Create a data frame for ggplot
df <- data.frame(x = x, pdf_normal = pdf_normal)

# Plot the normal distribution with shaded area from -3 to 0.5
ggplot(df, aes(x = x, y = pdf_normal))+
  geom_line(color = "blue", size = 1.5) +
  geom_ribbon(data = subset(df, x >= -3 & x <= 1.96),
              aes(ymax = pdf_normal, ymin = 0, fill = "Shaded Area"),
              alpha = 0.3, color = NA) +
  labs(title = "Normal(0,1)",
       x = "X",
       y = "Probability Density") +
  scale_fill_manual(values = c("Shaded Area" = "grey10"), guide = "none") +
  annotate("point", x = 1.96, y = 0, color = "green", size = 4) +
  annotate("text", x = 0, y = 0.15, label = "95%", size = 10, color = "red") +
  theme_minimal()+
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
# TODO: find m with qnorm
```

(b) Given $X \sim \text{N}(2,0.7^2)$, find the probability of $X$ being between 1 and 4.

```{r}
# TODO: calculate the probability
```

(c) Find the probability $P(5\leq X\leq 12)$ for $X \sim \text{N}(10,5)$.

```{r}
# TODO: calculate the probability
```

## 3. Approximating the Binomial with the Normal distribution

Please write down the questions and answers yourself, following the format above.
